---
title: "Practical Machine Learning"
author: "daisuke ohnuki"
date: "1/28/2021"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Practical Machine Learning Course Project

## Executive Summary
Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement – a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it.
The goal of this project is to predict the manner in which they did the exercise. 

## Data Set
```{r}
library(lattice); library(ggplot2);library(caret) ;library(rpart) ;library(rpart.plot)
library(RColorBrewer) ;library(rattle) ;library(randomForest)
set.seed(1989)
```

```{r}
trainUrl <- "http://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
training <- read.csv(url(trainUrl), na.strings=c("NA","#DIV/0!",""))
testUrl <- "http://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
testing <- read.csv(url(testUrl), na.strings=c("NA","#DIV/0!",""))
```
```{r , echo=FALSE}
getwd()
```

## Data Cleaning
```{r}
NAindex <- apply(training,2,function(x) {sum(is.na(x))}) 
training <- training[,which(NAindex == 0)]
NAindex <- apply(testing,2,function(x) {sum(is.na(x))}) 
testing <- testing[,which(NAindex == 0)]
```
### Cleaning the variables
```{r}
v <- which(lapply(training, class) %in% "numeric")

preObj <-preProcess(training[,v],method=c('knnImpute', 'center', 'scale'))
trainLess1 <- predict(preObj, training[,v])
trainLess1$classe <- training$classe

testLess1 <-predict(preObj,testing[,v])
```

### Remove non zero variables
```{r}
nzv <- nearZeroVar(trainLess1,saveMetrics=TRUE)
trainLess1 <- trainLess1[,nzv$nzv==FALSE]
nzv <- nearZeroVar(testLess1,saveMetrics=TRUE)
testLess1 <- testLess1[,nzv$nzv==FALSE]
```

```{r}
set.seed(1989)
inTrain= createDataPartition(trainLess1$classe, p = .75, list=FALSE)
training = trainLess1[inTrain,]
crossValidation = trainLess1[-inTrain,]
```

```{r}
modFit <- train(classe~., method="rf", data=train, trControl=trainControl(method='cv'), number=5, allowParallel=TRUE )
```

```{r}
trainingPred <- predict(modFit, training)
confusionMatrix(trainingPred, training$classe)
```



## Decision Tree approach
```{r}
modFital = rpart(classe ~., data=training, method ="class")
fancyRpartPlot(modFital)
```

## Random Forest approach
```{r}
modFitb1 <- randomForest(classe ~., data=training)
predictionB1 <- predic(modFitb1, training , type ="class")

```

```{r}
importance(modFitb1)
```
```{r}
plot(randomForest(classe ~., data=training))
```
Seeing the analysis above, Random Forest approach gives better qualification.

## Applying to prediction
```{r}
predictiionAL <- predic(modiFItAl, testing, type = "class")
confusionMatrix(predictionAL, training$classe)
```


## References
Velloso, E.; Bulling, A.; Gellersen, H.; Ugulino, W.; Fuks, H. Qualitative Activity Recognition of Weight Lifting Exercises. Proceedings of 4th International Conference in Cooperation with SIGCHI (Augmented Human '13) . Stuttgart, Germany: ACM SIGCHI, 2013.

